|

### 2023-03-28 paper_comparison/paper_comparison/paper_comparison.Rmd
modify the x_dist text
update conclusion
remove some unused comments

(introduction refine - rank: low)

@@

The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two predictors. Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test. In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin. This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures. Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you. They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals. (expand?)

@@

(should the below p goes to conclusion as a discussion?)

@@

A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern. According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials. But the chance of detecting the "triple-U" shape drops significantly compared to other shapes. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four (@ramsey_tests_1969). However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape. The recommendation of the polynomial power for the RESET test should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.

@@

Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution. For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study]. And those additional predictors are linear transformation of random vectors following certain predictor distributions. What is truly unexpected is the huge variability in power of conventional tests compared to visual tests. The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively. The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level. This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation. It also suggests that, under the designed scenarios in the paper, the visual test is a more robust test.

@@

As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test. Structure is hardest to detect if fitted values are lognormal. Also, complex structure are generally harder to detect, but there are outliers. A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.

>>>>

@@

The effectiveness of lineup protocol for regression analysis is validated by @majumder_validation_2013 under relatively simple settings with up to two predictors. Their results suggest that visual tests are capable of testing the significance of a single predictor with a similar power as a t-test, though they express that in general it is unnecessary to use visual inference if there exists a conventional test, and they do not expect the visual test to perform equally well as the conventional test. In their third experiment, where there is not a conventional test, visual test outperforms the conventional test for a large margin. This is encouraging, as it promotes the use of visual inference in situations where there are no existing statistical testing procedures. Visual inference have also been integrated into diagnostic of hierarchical linear models by @loy2013diagnostic, @loy2014hlmdiag and @loy2015you. They use lineup protocols to judge the assumption of linearity, normality and constant error variance for both the level-1 and level-2 residuals.

@@

@@

A primary factor contributes to the non-linearity model is the shape of the non-linearity pattern. According to Figure \ref{fig:poly-power-uniform-j}, conventional tests have higher power in testing shapes constructed from lower orders of Hermite polynomials. But the chance of detecting the "triple-U" shape drops significantly compared to other shapes. To understand why this is, one needs to return to the way the RESET test is applied. It requires a parameter indicating degree of fitted values to test for, and the recommendation is to generically use four [@ramsey_tests_1969]. However, the "triple-U" shape constructed from the Hermite polynomials use power up to 18. If the RESET test had been applied using a higher power no less than six, the power curve of "triple-U" shape will be only slight lower than the power curve of "M" shape. The recommendation of the polynomial power for the RESET test should be revised, perhaps. This illustrates the sensitivity of conventional testing to the parameters, and it also points to a limitation that one needs to know the data generating process in order to set the parameters for the test.

@@

Although it is unusual to consider fitted value distribution or predictor distribution in power analysis, it is not so surprised that the power of conventional tests vary because of fitted value distribution. For instance, the power of the RESET test which is effectively a F test depends on the realizations of the additional predictors included in the auxiliary regression equation [@jamshidian2007study]. And those additional predictors are linear transformation of random vectors following certain predictor distributions. What is truly unexpected is the huge variability in power of conventional tests compared to visual tests. The discrete uniform distribution constantly produce the highest power while the normal distribution and the lognormal distribution produce the lowest power for non-linearity and heteroskedasticity pattern respectively. The difference in power corresponding to different distributions peak at around 90% when the effect size is at a moderate level. This huge difference is undesirable as is not uncommon for analysts to omit the factor of predictor distribution in power calculation and sample size calculation.

@@

As expected, if the fitted value distribution is not uniform, there is a loss of power in the visual test. Structure is hardest to detect if fitted values are lognormal. Also, complex structure are generally harder to detect, but there are outliers. Under the designed scenarios in this paper, we find the visual test to be a more robust test against the change of fitted value distributions. A surprising finding was that the direction of heteroskedasticity appears to affect the ability to visually detect it, with wedge to the right being less detectable.
